---
output:
  word_document: default
  html_document: default
---
```{r}
library(tidyverse)
library(tidymodels)
library(GGally)
library(lmtest)
```

```{r}
air <- airquality
```

```{r}
str(air)
summary(air)
```

```{r}
air2 <- drop_na(air)
str(air2)
```

```{r}
ggpairs(air2)
```
```{r}
ggcorr(air2,label=TRUE)
```

Build model

```{r}
lin_reg = recipe(Ozone ~ Temp, air2)

# Define the model
lm_model = linear_reg() 
lm_model = set_engine(lm_model, "lm") 

# Initiate the workflow
lm_wflow = workflow()

# Add the model to the workflow
lm_wflow = add_model(lm_wflow, lm_model)

# Add the recipe to the workflow
lm_wflow = add_recipe(lm_wflow, lin_reg)

# Fit the model
lm_fit = fit(lm_wflow, air2)

summary(lm_fit$fit$fit$fit)

```

Examine prediction interval

```{r}
#Prediction interval
temp_var = predict(lm_fit$fit$fit$fit, interval = "prediction") #accessing the fit object with the three $fit
new_df = cbind(air2, temp_var)

ggplot(new_df, aes(x = Temp, y = Ozone)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  geom_line(aes(y=lwr), color = "red", linetype = "dashed") +
  geom_line(aes(y=upr), color = "red", linetype = "dashed") +
  theme_bw() 
```
Get lower and upper bound of slope at 95% confidence interval

```{r}
# Assuming lm_fit is your fitted model object

# Load the required library
library(broom)

# Extract the model object from the workflow fit
lm_model_fit = lm_fit$fit$fit$fit

# Obtain a tidy summary of your model
tidy_summary <- tidy(lm_model_fit)

# Print the tidy summary
print(tidy_summary)

# To get the confidence interval for the slope you can do:
confint(lm_model_fit, 'Temp', level = 0.95)

```
Predict Y with X

```{r}
#Using predict function
predict_values = data.frame(Temp = c(80))
predict(lm_fit, new_data = predict_values)
```
Diagnostics

Examine scatterplot to test Assumption 1:

**Assumption 1:** The predictor and response variable have a linear relationship  

```{r}
ggplot(air2, aes(x=Temp,y=Ozone)) + geom_point() + 
  geom_smooth(method="lm",se=FALSE, color="red") + theme_bw()
```

Perform Durbin-Watson test to test independence of residuals for Assumption 2:

**Assumption 2:** Model errors (residuals) are independent  

```{r}
dwtest(lm_fit$fit$fit$fit)
```
We fail to reject the null hypothesis with a p-value greater than 0.05. This suggests that the residuals are likely independent.  

Examine a plot of residual to test Assumption 3:

**Assumption 3:** Model residuals exhibit constant variance  

```{r}
air2 = mutate(air2, resid1 = lm_fit$fit$fit$fit$residuals)
ggplot(air2,aes(x=Temp,y=resid1)) + geom_point() + theme_bw()
```
Looks good, no clustering

Finally we examine a histogram of residuals to check Assumption 4:

**Assumption 4:** Model residuals are Normally-distributed  

```{r}
ggplot(air2,aes(x=resid1)) + geom_histogram() + theme_bw()
```

Kind of, I see a right skew though.